Global Assessments of Fishery Status need Better Data more than Better Models

### 2020-10-15 documents/ovando-etal-assessing-global-fisheries-long.Rmd
waiting to hear back before merging to master (#7)

<!-- * more assertive on what the data needs really are -->
<!--   * where to apply globally  -->
<!-- * For places with less coverage, what would we actually recommend there  -->
<!-- * what do we know about catchabiliy -->
<!-- - e.g. 51, fleets became industrial  -->
<!--   - hammer down on that's the issue with the effort  -->
<!--   - how do we translate effort into f bey area -->
<!--   - here's  -->

@@

49% of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020]. The SOFIA process uses these formal assessments for their determinations of stock status wherever possible. However, that leaves 51% of global fisheries landings lacking in formal stock assessments. While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion. While local experts can be well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable (and expert opinion can also be misinformed). Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018;@thorson2012a].

Formal stock assessments require at a minimum robust timeseries of captures and a reliable index of abundance, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE). While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist [@maureaud2020], and would be prohibitively expensive to collect. Even when data are available though, running an individualized stock assessment is not a "just press play" process. Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment. This process takes considerable time and expertise.

Data-limited methods applied to the question of global stock status typically make compromises on both the data requirements and individual attention needed for formal stock assessment. Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (Free et al. 2020 [@free2020] and references therein). These models seek to infer stock status, for example in terms of biomass *B* relative to the biomass at maximum sustainable yield *B~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013]. In addition, a global assessment requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time. As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, global data-limited methods typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks. Each of the published catch-based methods reviewed in Free et al. 2020 [@free2020] sought to address specific gaps in our understanding of global fisheries, and in doing so introduced unique strengths and weaknesses.

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~. These issues become apparent when we consider some of the macro-level predictions made by these models. The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020]. A simple benchmark then is to compare the best available estimates of fishery status from RAM to those predicted by less reliable methods intended for use when insufficient data are available for a full stock assessment model.

Examining stock assessment data from RAM from the  Northeast Pacific,the Northeast Atlantic and the Mediterranean & Black Sea, the Northeast Pacific is in the best state (mean RAM B/B~MSY~ = 1.61), NE Atlantic (median RAM b/b~MSY~ =1.27) and the Mediterranean the worst (mean RAM B/B~MSY~ =0.52) [@hilborn2020]. The Costello et al. 2016 [@costello2016a] model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison. The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or NOrtheast Pacific. The catch based approach @pauly2007, which compares catch in the last year to maximum catch and equates these to stock status, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)). While the methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks,

@@

The coming decades are a critical period for the future of global oceans. Climate change, population growth, and economic expansion will all place increasing pressure on fisheries. While our understanding of global fisheries has improved greatly,  mitigating these threats, as called for by SDG 14, will require an improved ability to track the status of fisheries around the world, particularly those stocks that are currently unassessed. At the same time, new sets of globally distributed data have recently become available, along with advances in statistical modeling methods. We leverage these opportunities to provide a global assessment of our perception of global stock status resulting from different use of different data sources. Our results demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection and interpretation effort.

@@

@@

For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index and swept area ratio scores. We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock. We used these data to demonstrate how `sraplus` can add incremental amounts of information to an assessment, and the potential value of doing so. We first used the catch history heuristics internal to CMSY [@froese2017] to estimate stock status. We then used stock-specific data on swept area ratio and fisheries management index to generate priors on F/F~MSY~ for each of the stocks, which were then passed to `sraplus`. It is critical to note that in both of these cases the catch data on its own contains no information about stock status; we can only produce estimates of stock status by combining catch data with explicit priors and specified functional relationships [@thorson2013]. This is important since it means that we must look to our priors and functional forms to understand where our insight stock status is coming from, not the catch data alone. Lastly, we used the reconstructed effort data from @rousseau2019 to create an index of abundance for each stock (see [Estimation Model](#Estimation Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.

Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one. CMSY, using no stock-specific information but the catch history and the estimated resilience of the target species, badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing. Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~. Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit. This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values. An abundance index based on catch and regional effort cannot capture this trend, since it interprets the extremely low catches (near zero) as low biomass since we do not have sufficiently fine-scale effort data to reflect the reduction in effort that goes hand-in-hand with regulated catch reductions (Fig.\@ref(fig:cs-plot)).

@@

That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise. In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process. Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status.

@@

As our test case we use `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM with estimates of B/B~MSY~ and greater than 25 years of continuous catch history. B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then pair the catch histories for these stocks with regional-level swept area ratio, fisheries management index, and effort data. For any given stock we use the area-weighted swept area ratio (where area refers to the spatial area of the SAR study in question) for the FAO region of the stock. For fisheries management index we use the mean component FMI values for the species group and FAO region in question (or FAO region if no matches exist at the species group level). This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries. We then constructed an effort index for the stock in question as the total effort from @rousseau2019 estimated in all countries that report catch of the species in question within its FAO region, weighting the effort of each country by its contribution to the total catch of that species in that FAO region.

@@

```{r dat-desc}
dat_description <- tribble(~`Data Name`, ~Description,
        "RAM-Data", "Fit to abundance index from RAM",
        "sar", "Prior on terminal F/Fmsy set by regional swept area ratio",
        "fmi", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
        "cpue", "Fit to CPUE index created from RAM catch and regional effort index. 2.5% tech. creep",
        "cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.5% tech. creep",
        "nominal-cpue", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
         "nominal-cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
        "catch-only", "priors on terminal B/Bmsy informed by catch history",
        "heuristic", "priors on initial and terminal B/Bmsy informed by Cmsy heuristc")

@@

Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries. Since no global database of length or age composition data exists, global assessment methods are constrained to surplus production models in the manner used by `sraplus`. This creates two main reasons why `sraplus` might produce poor results at a global level: model misspecification and poor data quality. Model misspecification errors might result from the surplus production model underlying `sraplus` being fundamentally unable to capture the more complicated dynamics resulting from the full age-structured population models common in RAM. Errors resulting from poor quality data mean that even if the underlying model is a reasonable approximation of the real world, the model may produce incorrect results due to the data supplied to the model simply not containing accurate information on the phenomenon in question.

Assessing stock status through `sraplus` using the catch history and an index of abundance from RAM gives us a sense for how much model miss-specification might limit the ability of global assessment relative to data quality. Overall the `sraplus` estimates using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy  = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). This exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.

<!-- The median percent error (a measure of bias) in B/B~MSY~ for the most recent year that data was available of models fit to the RAM data was `r percent(perf$mpe[perf$data == "ram-data"])`, indicating a slight positive bias. The mean absolute percent error was `r percent(perf$mape[perf$data == "ram-data"])`, and the model correctly classified stocks to the FAO stock status bin `r percent(perf$accuracy[perf$data == "ram-data"])` of the time. Looking across space, we see clear geographic patterns in model performance. The RAM-data based model estimates performed extremely well in FAO region 67 (encompassing the waters of Alaska and western Canada). The model struggled the most in region XX, comprised of stocks XX (some thought on why) (Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). While these results are not perfect, this exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably able to reproduce the global state of fisheries as understood from formally assessed fisheries.  -->

Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves. These becomes clearer once we consider the performance of `sraplus` models fit to combinations of our broadly available datasets. Many of the datasets used produced similar levels of bias as the RAM data (Tab.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The constraints of population dynamics mean that B/B~MSY~ is usually on the range of 0 to 4. The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near 1. This means that a model that more or less reproduces the global average of stock status (near 1) will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])` (since fisheries are not uniformly distributed across each bin).

Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process. Across the models performance (or lack thereof) is not consistent in space: use of different data performed best or worst for different FAO regions. For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map). We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)). Overall, while some data sources performed slightly better than others by some metrics in some places, there was no consistent spatial pattern to this performance, and no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.

@@

Global-level assessment are critical for guiding management agendas for the world's oceans. Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world. While in some cases addition of globally available data such as quality of fisheries management or effort reconstructions provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).

What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends of course on the needs of those using the assessment outputs. It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient. In some instances using the data presented here did provide some improvement over use of catch-only style methods; the difficulty comes in attempting to apply data types uniformly across the globe. While it is unreasonable to expect any global-scale data to be able to perform as well as data pulled from RAM assessments themselves, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general trends in stock status in time and space. That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the data themselves.

@@

Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions. The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e. those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)). Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet. Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`. But, applying one SAR value to all stocks in a region causes inaccurate estimates of stock status. Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.

It is easy enough to call for expanded data collection at a global scale. In world of finite resources though, it is important to consider what sources of data might provide the greatest value in improving our estimates of global stock status. We fit 3,000 `sraplus` models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated. We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).

Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates in the most recent five years by on average 15%. Information on fishing mortality rates (F/F~MSY~) in the most recent year was equally useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ estimates was actually more useful Fig.\@ref(fig:voi-plot). Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance. While having access to a complete timeseries of an abundance index had a higher probability of reducing RMSE than increasing it, in many cases having only an abundance index resulted in a correct fit to the overall trend in abundance, but an error-prone estimate of B/B~MSY~ itself due to incorrectly estimating either the shape parameter, the initial state of the stock, or both. However, using an index of abundance that only covered the most recent quarter of a fishery's history actually increased RMSE by over 10% on average.

@@

\begin{equation}
  p_t \sim e^{normal\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)}
  (\#eq:procerror)
\end{equation}

>>>>

<!-- * more assertive on what the data needs really are -->

<!--   * where to apply globally  -->

<!-- * For places with less coverage, what would we actually recommend there  -->

<!-- * what do we know about catchabiliy -->

<!-- - e.g. 51, fleets became industrial  -->

<!--   - hammer down on that's the issue with the effort  -->

<!--   - how do we translate effort into f bey area -->

<!--   - here's  -->

@@

Forty nine percent of landings of marine fishes reported by the FAO are represented in the RAM Legacy Stock Assessment Database, a repository of "gold standard" estimates of fishery status [@hilborn2020]. The SOFIA process uses these formal assessments for their determinations of stock status wherever possible. However, that leaves 51% of global fisheries landings lacking in formal stock assessments. While these "unassessed" stocks are generally individually smaller than the typically larger and more valuable stocks in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report bases its estimates of included unassessed stocks mostly on expert opinion. While local experts can be well informed as to the status of their fish stocks, a more quantitative and reproducible process would be desirable (and expert opinion can also be misinformed). Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide estimates for the global status of these unassessed stocks [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a].

Formal stock assessments require at a minimum robust timeseries of captures and a reliable index of abundance and/or length composition data, ideally coming from scientifically designed surveys but sometimes from fishery dependent catch-per-unit-effort (CPUE). While most of the world's largest and most valuable fisheries have these data and have stock assessments based on the data [@hilborn2020], for many of the world's fisheries these data do not exist [@maureaud2020], and would be prohibitively expensive to collect. Even when data are available though, running an individualized stock assessment is not a "just press play" process. Assessment scientists must make careful considerations about the quality and subsequent weighting of available data, the level of biological abstraction that the species and the data can support in the assessment model, and of course the quality of results produced by the assessment. This process takes considerable time and expertise.

When applied in bulk to the question of global stock status, data-limited assessment methods typically make compromises on both the data requirements and individual attention needed for formal stock assessment. Given that catch statistics are, to date, the only globally available form of comparable fishery data, global data-limited assessments have typically been forms of "catch-only" models (Free et al. 2020 [@free2020] and references therein). These models seek to infer stock status, for example in terms of biomass *B* relative to the biomass at maximum sustainable yield *B~MSY~*, from characteristics of a fishery's catch history, for example the ratio of catch to maximum catch [@martell2013]. In addition, a global assessment requires producing results for hundreds or thousands of individual fisheries within a reasonable amount of time. As such, rather than spending time carefully evaluating the quality of the data, model assumptions, and model fit for individual stocks, data-limited methods applied across thousands of stocks at the global level typically aim for a method that is on average correct, but may not be reliable for assessment of individual stocks. Each of the published catch-based methods reviewed in Free et al. 2020 [@free2020] sought to address specific gaps in our understanding of global fisheries, and in doing so offer unique strengths and weaknesses.

However, Free et al. 2020 [@free2020] demonstrated that these catch-only models can often produce both imprecise and biased estimates of current stock status in terms of B/B~MSY~. These issues become apparent when we consider some of the macro-level predictions made by these models. The RAM Legacy Stock Assessment Database [@ricard2012] contains the best available estimates of B/B~MSY~ and other fishery reference points for hundreds of fisheries making up 49% of global marine catches [@hilborn2020]. While it must noted that the estimates in RAM are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the best available estimates of fishery status from RAM to those predicted by potentially less reliable methods intended for use when insufficient data are available for a full stock assessment model.

Examining stock assessment data from RAM from the Northeast Pacific,the Northeast Atlantic and the Mediterranean & Black Sea, the Northeast Pacific is in the best state (mean RAM B/B~MSY~ = 1.61), NE Atlantic (median RAM b/b~MSY~ =1.27) and the Mediterranean the worst (mean RAM B/B~MSY~ =0.52) [@hilborn2020]. The Costello et al. 2016 [@costello2016a] model shows the general pattern most fishery scientists could agree on, but the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be over-optimistic, and the Northeast Pacific should be better by comparison. The Rosenberg super-ensemble method [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or NOrtheast Pacific. The catch based approach [@pauly2007], which compares catch in the last year to maximum catch and equates these to stock status, finds the stocks of Southeast Asia in much better condition than the Northeast Pacific or Northeast Atlantic, in contrast to the general perception (Table.\@ref(tab:status-tab)). The methods besides RAM in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the NOrtheast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.

@@

The coming decades are a critical period for the future of global oceans. Climate change, population growth, and economic expansion will all place increasing pressure on fisheries. While our understanding of global fisheries has improved greatly, mitigating these threats, as called for by SDG 14, will require an improved ability to track the status of fisheries around the world, particularly those stocks that are currently unassessed. At the same time, new sets of globally distributed data have recently become available, along with advances in statistical modeling methods. We leverage these opportunities to show how our perception of global stock status changes depending on the data source used in an assessment. Our results demonstrate that modeling improvements alone cannot provide a satisfactory answer to the question of global fishery status, requiring instead a robust and worldwide data collection and interpretation effort.

@@

+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Data Source                                 | Short Name | Data Use                                                                 | Caveats                                                                                                         |
+=============================================+============+==========================================================================+=================================================================================================================+
| Catch data [@fao2020]                       | catches    | Priors on stock status, scaling of population size, exploitation history | Heuristics or regressions used to translate shape of catch history into priors on stock status                  |
+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Fisheries Management Index [@melnychuk2017] | FMI        | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Swept Area Ratio [@amoroso2018]             | SAR        | Priors on most recent F/F~MSY~ values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Reconstructed effort data[@rousseau2019b]   | effort     | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors. Assumed rate of technology creep reported in individual sections |
+---------------------------------------------+------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+

: Broadly available data used to inform estimates of global stock status

@@

For our case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific Fisheries Management Index (a measure of management capacity) and swept area ratio (a measure of trawl fishing pressure) scores. We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al. 2019 [@rousseau2019] to each stock. We used these data to demonstrate how `sraplus` can add incremental amounts of information to an assessment, and the potential value of doing so. We first used the catch history heuristics internal to CMSY [@froese2017] to estimate stock status, due to its widespread use [XX do we really want to cite those papers...XX]. We then used stock-specific data on swept area ratio and fisheries management index to generate priors on F/F~MSY~ for each of the stocks, which were then passed to `sraplus`. Lastly, we used the reconstructed effort data from [@rousseau2019] to create an index of abundance for each stock (see [Estimation Model](#Estimation%20Model) section), and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's fisheries management index and swept area ratio values.

Nearly all of the fisheries used in this case study have F/F~MSY~ values less than one, and most have B/B~MSY~ values greater than one. CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of B/B\~msy\~) and the estimated resilience of the target species (based on Fishlife ([@thorson2020])), badly misses this trend, stating instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing. Adding priors informed by stock-specific the swept-area-ratio and fisheries management index data produces a dramatically different picture, capturing the general state that most of these fisheries have low fishing mortality rates and biomass near or above B~MSY~. Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit. This is largely because a number of stocks in this case study set have very low catches (relative to historic highs) and high biomass values (Fig.\@ref(fig:cs-plot)).

@@

<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise. In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process. Since the SAR and FMI regressions were trained on the RAM stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status. -->

@@

Moving beyond our case study, our goal is to determine the ability of FMI, SAR, or effort data to improve estimates of global stock status. We based this test on `r ram_comp_data$stockid %>% n_distinct()` fisheries from RAM, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of B/B~MSY~ and greater than 25 years of continuous catch history. B/B~MSY~ values from RAM are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RAM stocks with regional-level swept area ratio, fisheries management index, and effort data. For any given stock we use the area-weighted swept area ratio (where area refers to the spatial area of the SAR study in question) for the FAO region of the stock. For fisheries management index we use the mean component FMI values for the species group and FAO region in question (or FAO region if no matches exist at the species group level). This process approximates a global-level assessment exercise, where data are available at regional levels, but not for specific fisheries. We then constructed an effort index for the stock in question as the total effort from Rousseau et al. 2019 [@rousseau2019] estimated in all countries that report catch of the species in question within its FAO region, weighting the effort of each country by its contribution to the total catch of that species in that FAO region.

@@

```{r dat-desc}
dat_description <- tribble(~`Data Name`, ~Description,
        "RAM-Data", "Fit to abundance index from RAM",
        "sar", "Prior on terminal F/Fmsy set by regional swept area ratio",
        "fmi", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
        "cpue", "Fit to CPUE index created from RAM catch and regional effort index. 2.5% tech. creep",
        "cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 2.5% tech. creep",
        "nominal-cpue", "Fit to CPUE index created from RAM catch and regional effort index. 0% tech. creep",
         "nominal-cpue-plus", "Fit to CPUE index created from RAM catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
        "catch-only", "priors on terminal B/Bmsy informed by catch history",
        "heuristic", "priors on initial and terminal B/Bmsy informed by CMSY heuristc")

@@

Stock assessment models included in RAM are generally age structured models containing a variety of nuances specific to individual species and fisheries. Since no global database of length or age composition data exists, most global assessment methods use surplus production models in the manner used by `sraplus`. This creates two main reasons why `sraplus` might produce poor results at a global level:  first, model misspecification errors might result from the surplus production model underlying `sraplus` being fundamentally unable to capture the more complicated dynamics resulting from the full age-structured population models common in RAM. Second, errors resulting from poor quality data might mean that even if the underlying model is a reasonable approximation of the real world, the model may produce incorrect results due to the data supplied to the model simply not containing accurate information on the phenomenon in question.

Overall the `sraplus` estimates of B/B~MSY~  resulting from using the RAM data are reasonably good (median absolute percent error `r percent(perf$mape[perf$data == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$data == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). This exercise tells us that given sufficiently high quality data, a surplus production model such as `sraplus` is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.

Performance limitations then are likely to arise less from model misspecification than from the quality of the data themselves. These becomes clearer once we consider the performance of `sraplus` models fit to combinations of our broadly available datasets. Many of the datasets used produced similar levels of bias as the RAM data (Tab.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The constraints of population dynamics mean that B/B~MSY~ is usually on the range of 0 to 4. The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near 1. This means that a model that more or less reproduces the global average of stock status (near 1) will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$data %in% c("ram-data","guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category has a mean accuracy of `r percent(perf$accuracy[perf$data == "guess"])`.

Looking geographically we see a similar pattern of a rapid decrease in performance for models using non-RAM data intended to simulate a global assessment process. Across the models, performance (or lack thereof) was not consistent in space: use of different data performed best or worst for different FAO regions. For example, models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) Fig.\@ref(fig:mape-map). We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)). Overall, while some data sources performed slightly better than others by some metrics in some places, there was no consistent spatial pattern to this performance, and no models using any non-RAM data were able to capture the overall state or geographic distribution of stock status represented in RAM in a consistently satisfactory manner.

@@

Global-level assessment are critical for guiding management agendas for the world's oceans, and tracking critical indicators such as the United Nations Sustainable Development Goals. Despite this need, and despite advances in stock assessment methods and available data, we show that our understanding of the world's fisheries remains murky in many parts of the world. While in some cases addition of globally available data, such as quality of fisheries management or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RAM-derived indices, produced biased and imprecise estimates of stock status, frequently performing worse than a simple guess (Table.\@ref(tab:perf-tab)).

What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality dependson the needs of those using the assessment outputs. It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficient. In some instances using the data presented here did provide some improvement over use of catch-only style methods; the difficulty comes in attempting to apply data types uniformly across the globe. While it is unreasonable to expect any global-scale data to be able to perform as well as data pulled from RAM assessments themselves, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general trends in stock status in time and space. That none of the datasets collected here can achieve that, and that our test on the RAM data suggest that model misspecification is not the primary culprit, tells us that improvements in estimates of global stock status must come from improvements in the data themselves.

@@

Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions. The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e. those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)). Effort data such as those reconstructed by Rousseau et al. 2019 [@rousseau2019] can help distinguish between regions with similar catch histories but different effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet. Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of F/F~MSY~ of any of the datasets we explore on an individual stock basis, a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)`.

But, we must simulataneously consider the ideas of data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly. Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources.

What sources of data might provide the greatest value in improving our estimates of global stock status? We fit 3,000 `sraplus` models to randomly sampled fisheries from RAM, each time varying the kind and quality of data made available to the model, and what parameters the model estimated. We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)).

Having knowledge of the state of the population in the most recent year reduced RMSE of the model estimates by on average 15%. Information on fishing mortality rates (F/F~MSY~) in the most recent year was equally useful as data on B/B~MSY~ itself, and having access to a recent string of estimates of F/F~MSY~ (which is conditional on the recent fishing selectivity regime) was actually more useful (Fig.\@ref(fig:voi-plot)). Interestingly, having access to an estimate of F/F~MSY~ had a clearer and more substantial effect than having an index of abundance. While having access to a complete timeseries of an abundance index had a higher probability of reducing RMSE than increasing it, in many cases having only an abundance index resulted in a correct fit to the overall trend in abundance, but an error-prone estimate of B/B~MSY~ itself due to incorrectly estimating either the shape parameter, the initial state of the stock, or both. However, using an index of abundance that only covered the most recent quarter of a fishery's history actually increased RMSE by over 10% on average.

@@

\\begin{equation} p\_t \\sim e\^{normal\\left(-\\sigma\_{proc}\^2/2,\\sigma\_{proc}\\right)} (\\\#eq:procerror) \\end{equation}
